Sting nematodes modify metabolomic profiles of host plants

### 2019-10-18 manuscript/sci-rep/reviewer-letter/reviewer-response.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{changepage}% http://ctan.org/pkg/changepage
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{ragged2e}
\usepackage{scrextend}
\usepackage{lastpage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\rhead{\includegraphics[height= 1cm]{ento1.png}\vspace{1em}}
\chead{\includegraphics[width = 5cm]{agritech1.png}}
\lhead{\includegraphics[width=3cm]{acet1.png}}

\fancyfoot{}
\setlength\headheight{40.8pt}
\RequirePackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{titlesec}
\titleformat{\section}{\normalfont\fontsize{10}{15}\bfseries}{\thesection}{0em}{}
\titlespacing{\section}{}{1.24em}{0.24em}

\cfoot{Page \thepage \hspace{1pt} of \pageref{LastPage}}

\begin{document}

\begin{addmargin}[2.8in]{}
Willett, Filgueiras, \newline
Benda, Zhang, and Kenworthy \newline
----------------------------- \newline
Denis S. Willett\newline
Department of Entomology \newline
Cornell AgriTech  \newline
Cornell University \newline
15 Castle Creek Drive \newline
Geneva, NY 14456 \newline
deniswillett@cornell.edu \newline
\end{addmargin}
\setlength{\parindent}{0cm}

October 24, 2019

\vspace{1.24em}

Scientific Reports

\vspace{1.24em}

Dear Dr. Geary,

\vspace{0.48em}
\setlength{\parindent}{1.24cm}

Thank you for your consideration of \textit{“Sting nematodes modify metabolomic profiles of host plants”} for publication in Scientific Reports.  We have revised the manuscript in accordance with reviewer recommendations.  A complete and detailed account of the revisions is included below.  The reviewer comments were very helpful, much appreciated, and have resulted in a better manuscript.  In addition to submitting a revised version of the manuscript incorporating the latest changes, we are also submitting production quality images.

\vspace{2em}

Much appreciated, \newline

\vspace{1em}

Denis S Willett
\vspace{0.48em}

Camila Cramer Filgueiras
\vspace{0.48em}

Nicole D Benda
\vspace{0.48em}

Jing Zhang
\vspace{0.48em}

Kevin E Kenworthy

\hline

\paragraph{Reviewer 1 (Remarks to the Author)}

You spend a lot of time talking about relative tolerance of the bermudagrass lines to sting nematode. However, I was never very clear on how you are defining tolerance or how this was determined.

On page 2 under Bermudagrass response to sting nematode feeding...I think you should use the term nematode population densities instead of "nematode densities" or "differences in nematode population."

Very well written and very interesting paper

Reviewer #2 (Technical Comments to the Author):

This is a well researched study and the manuscript well structured and written. The methods used are scientifically sound, results appropriately interpreted and discussed. Except for a few minor comments below, I recommend publication of this manuscript.

Line 58. The authors stated that the canonical correspondence analysis explained 42\% of the observed inertia in the total metabolome. I suggest that the authors explain this because in Figure 2 this analysis explained 71.9\% of the total metabolome variation (CCA1: 45.3\% + CCA2: 26.6\%).

Lines 146 and 150. Change ‘conetainers’ to ‘containers and container’

I suggest that instead of putting in the appendix, the package names used for their statistical analysis, the authors should include these names in the text to educate readers, especially for those not familiar with the ‘R software’

Reviewer #2 (Remarks to the Author):

This is a well researched study and the manuscript well structured and written. The methods used are scientifically sound, results appropriately interpreted and discussed. Except for a few minor comments below, I recommend publication of this manuscript.

Line 58. The authors stated that the canonical correspondence analysis explained 42\% of the observed inertia in the total metabolome. I suggest that the authors explain this because in Figure 2 this analysis explained 71.9 \% of the total metabolome variation (CCA1: 45.3\% + CCA2: 26.6\%).

Lines 146 and 150. Change ‘conetainers’ to ‘containers and container’

I suggest that instead of putting in the appendix, the package names used for their statistical analysis, the authors should include these names in the text to educate readers, especially for those not familiar with the ‘R software

\end{document}

### 2019-10-18 manuscript/sci-rep/reviewer-letter/reviewer-response.tex

Kevin E Kenworthy

\vspace{1.4em}

\hline

\setlength\parindent{0pt}

\subsection*{Reviewer 1 (Remarks to the Author)}

@@

You spend a lot of time talking about relative tolerance of the bermudagrass lines to sting nematode. However, I was never very clear on how you are defining tolerance or how this was determined.

\begin{quote}
    \textit{This is an excellent question.  We have taken the opportunity to elaborate on this in the introduction.  In short, tolerance in bermudagrass is determined if there was no reduction in root length or had greater root length than ‘Tifway’ despite \textit{B. longicaudatus} infection.  Tolerance of the varieties had been established in previous work by Pang et al. 2011, citations 23, and 24 in the manuscript.  } 
\end{quote}

@@

On page 2 under Bermudagrass response to sting nematode feeding...I think you should use the term nematode population densities instead of "nematode densities" or "differences in nematode population."

\begin{quote}
    \textit{Thanks for pointing this out.  We have adjusted the language in that section in accordance with your suggestions.} 
\end{quote}

@@

Very well written and very interesting paper

\begin{quote}
    \textit{Thank you for your interest and constructive feedback.}
\end{quote}

\subsection*{Reviewer #2 (Technical Comments to the Author)}

@@

This is a well researched study and the manuscript well structured and written. The methods used are scientifically sound, results appropriately interpreted and discussed. Except for a few minor comments below, I recommend publication of this manuscript.

\begin{quote}
    \textit{Thank you for your feedback.  We appreciate the interest and comments below.}
\end{quote}

@@

Line 58. The authors stated that the canonical correspondence analysis explained 42\% of the observed inertia in the total metabolome. I suggest that the authors explain this because in Figure 2 this analysis explained 71.9\% of the total metabolome variation (CCA1: 45.3\% + CCA2: 26.6\%).

\begin{quote}
  \textit{Thanks for your inquiry.  Canonical correspondence analysis did explain 42\% of the observed inertia.  Of that 42\% explained inertia, axis 1 explained 45.3\% and axis 2 26.6\%.  Summing the explained variation of the other 3 axis with the first two would result in 100\% of the variation captured in canonical ordination or 42\% of observed inertia (constrained $\chi^2$) in the total metabolome. This explanation has been elaborated upon in the text.   } 
\end{quote}

@@

Lines 146 and 150. Change ‘conetainers’ to ‘containers and container’

\begin{quote}
    \textit{Thanks for suggesting this.  We may have been carried away with cones.  The text has been altered to read container on the lines mentioned.  }
\end{quote}

@@

I suggest that instead of putting in the appendix, the package names used for their statistical analysis, the authors should include these names in the text to educate readers, especially for those not familiar with the ‘R software’

\begin{quote}
    \textit{Thanks for your suggestion.  This is a good idea; we have incorporated names of the software packages in the main body of the text.}
\end{quote}

### 2019-10-18 manuscript/sci-rep/nature-plants.tex
Sci-rep cover letter
reviewer response draft and revised manuscript

To investigate whether sting nematodes modify metabolite production in their host plants, mixed age \textit{Belonolaimus longicaudatus} nematodes were introduced to the root zones of three African bermudagrass (\textit{Cynodon transvaalensis}) lines of differing tolerance: one susceptible line (AB03), one moderately tolerant line (AB33), and one tolerant line (AB39).  Comparison control plants received no nematodes.  Ninety days after inoculation,  Plants were unearthed and sting nematode damage assessed.  Following damage assessment, untargeted metabolomic profiling was conducted on root samples.

\subsection*{Bermudagrass response to sting nematode feeding}
Sting nematode densities were significantly ($df = 36$, $t > 4$, $p < 0.004$, Tukey's method) higher on moderately tolerant (AB33) bermudagrass (Figure \ref{fig:figure1}A).  (No sting nematodes were recovered from comparison plants not inoculated with sting nematodes; there was no contamination.)  No significant ($df = 36$, $t = -0.75$, $p = 0.73$) differences in sting nematode populations on susceptible (AB03) or tolerant (AB39) lines were observed.

@@

@@

To evaluate the effects of nematode infection on bermudagrass metabolite response, treated plants in conetainers were inoculated with 50 sting nematodes of mixed age and gender per conetainer.  Control (noninoculated) plants did not receive any nematode treatment. Seven replications of each line (AB03, AB33, AB39) and treatment (inoculated, noninoculated) combination were conducted.  Ninety days following nematode inoculation (time enough for nematode growth and reproduction), plants were removed from the conetainer, the soil plug removed, and the 100ml of sand gently rinsed to extract the nematodes for further processing through centrifugal flotation and counting on an inverted microscope.  Rinsed roots were placed in 50ml falcon tubes then immersed in liquid nitrogen and stored at -80\textsuperscript{o}C until lyophilization.  Lyophilzed roots were weighed and samples selected for metabolomics analysis.

@@

Raw LC/mass spectrometry data were uploaded to Metabolomics Workbench \cite{sud2015metabolomics}.  All analysis on the raw data was conducted in R version 3.5.2 using RStudio as an IDE (with Vim keybindings) \cite{rcore2018,rstudio}.  A full list of packages used to facilitate the analysis along with citations can be found in Appendix E.  All code, including manuscript documentation, is available on GitHub (https://github.com/acetworld/bermuda-grass-metabolomics).

@@

Must include all authors, identified by initials, for example:
K.E.K. conceived the experiments, N.D.B and J.Z. conducted the experiments, D.S.W. and C.C.F analyzed the results and wrote the manuscript.  All authors reviewed the manuscript.

\begin{figure}
\begin{fullwidth}
\includegraphics[width=0.95\linewidth]{figures/publication_figures/figure-2.pdf}
\caption{Differences in global metabolome of three bermudagrass lines inoculated and noninoculated with sting nematode.  Axes represent first two orthogonal axes from canonical correspondence analysis with percentages indicating relative contribution of each axis to constrained $\chi^2$.  Points represent the metabolome of individual bermugrass replicates projected into ordination space.  Ellipses are 95\% confidence intervals for each treatment-line combination.  Blue arrows represent explanatory variables; direction of arrow depicts direction of the gradient while length denotes relative (scaled) importance.  }
\label{fig:figure2}
\end{fullwidth}
\end{figure}

>>>>

To investigate whether sting nematodes modify metabolite production in their host plants, mixed age \textit{Belonolaimus longicaudatus} nematodes were introduced to the root zones of three African bermudagrass (\textit{Cynodon transvaalensis}) lines of differing tolerance: one susceptible line (AB03), one moderately tolerant line (AB33), and one tolerant line (AB39).  Tolerance in these bermudagrass lines had been determined in previous work; lines are considered tolerant if there was no reduction in root length or had greater root length than the ‘Tifway’ cultivar despite \textit{B. longicaudatus} infection \cite{pang2011bermudagrass, pang2011screening}.  In our work, comparison control plants received no nematodes.  Ninety days after inoculation,  Plants were unearthed and sting nematode damage assessed.  Following damage assessment, untargeted metabolomic profiling was conducted on root samples.

\subsection*{Bermudagrass response to sting nematode feeding}
Sting nematode population densities were significantly ($df = 36$, $t > 4$, $p < 0.004$, Tukey's method) higher on moderately tolerant (AB33) bermudagrass (Figure \ref{fig:figure1}A).  (No sting nematodes were recovered from comparison plants not inoculated with sting nematodes; there was no contamination.)  No significant ($df = 36$, $t = -0.75$, $p = 0.73$) differences in sting nematode population densities on susceptible (AB03) or tolerant (AB39) lines were observed.

@@

\begin{figure}
\begin{fullwidth}
\includegraphics[width=0.95\linewidth]{figures/publication_figures/figure-2.pdf}
\caption{Differences in global metabolome of three bermudagrass lines inoculated and noninoculated with sting nematode.  Axes represent first two orthogonal axes from canonical correspondence analysis with percentages indicating relative contribution of each axis to constrained $\chi^2$ (or 42\% of observed inertia of the global metabolome).  Points represent the metabolome of individual bermugrass replicates projected into ordination space.  Ellipses are 95\% confidence intervals for each treatment-line combination.  Blue arrows represent explanatory variables; direction of arrow depicts direction of the gradient while length denotes relative (scaled) importance.  }
\label{fig:figure2}
\end{fullwidth}
\end{figure}

@@

To evaluate the effects of nematode infection on bermudagrass metabolite response, treated plants in containers were inoculated with 50 sting nematodes of mixed age and gender per container.  Control (noninoculated) plants did not receive any nematode treatment. Seven replications of each line (AB03, AB33, AB39) and treatment (inoculated, noninoculated) combination were conducted.  Ninety days following nematode inoculation (time enough for nematode growth and reproduction), plants were removed from the container, the soil plug removed, and the 100ml of sand gently rinsed to extract the nematodes for further processing through centrifugal flotation and counting on an inverted microscope.  Rinsed roots were placed in 50ml falcon tubes then immersed in liquid nitrogen and stored at -80\textsuperscript{o}C until lyophilization.  Lyophilzed roots were weighed and samples selected for metabolomics analysis.

@@

Raw LC/mass spectrometry data were uploaded to Metabolomics Workbench \cite{sud2015metabolomics}.  All analysis on the raw data was conducted in R version 3.5.2 using RStudio as an IDE (with Vim keybindings) \cite{rcore2018,rstudio}.  Packages used to facilitate analysis include: tidyverse \cite{tidyverse}, devtools \cite{devtools}, pacman \cite{pacman}, MASS \cite{mass}, Hmisc \cite{hmisc}, cowplot \cite{cowplot}, stringr \cite{stringr}, foreach \cite{foreach}, testthat \cite{testthat}, VIM \cite{VIM}, vsn \cite{vsn}, pvclust \cite{pvclust}, dendextend \cite{dendextend}, lettercase \cite{lettercase}, vegan \cite{vegan}, lmtest \cite{lmtest}, emmeans \cite{emmeans}, car \cite{car}, candisc \cite{candisc}, indicspecies \cite{indicspecies}, ggdendro \cite{ggdendro}.  All code, including manuscript documentation, is available on GitHub (https://github.com/acetworld/bermuda-grass-metabolomics).

@@

K.E.K. conceived the experiments, N.D.B and J.Z. conducted the experiments, D.S.W. and C.C.F analyzed the results and wrote the manuscript.  All authors reviewed the manuscript.
